Exploring the Usefulness of Varactor loaded Nonlinear Transmission Lines

### 2020-06-03 RohdeAndSchwarzWhitePaper/Draft_White_Paper.tex
Small random edits
Completed intro sections (second draft anyway) and fixed references

\section{ Theory }
\subsection{Why use Nonlinear Devices?}

@@

\subsection{Types of NLTL waves}
Solitons and Shockwaves.

\subsection{Varactors vs. Nonlinear Capacitors}

@@

\begin{equation}\label{eq:LargeSignalDiodeCap}
C_{ls}(V_{l},V_{n}) = \frac{1}{V_{n}-V_{l}}\int_{V_{l}}^{V_{n}}C(v)\, dv
\end{equation}

\begin{equation}\label{eq:bragg}
	f_{bragg} \approx \frac{1}{2\pi{}\sqrt{LC(v_{max})}}	
\end{equation}

\begin{equation}\label{eq:ShockwaveSections}
	n \approx \frac{T_{f,in}}{\Delta{}T}
\end{equation}

Where $T_{f,in}$ is the fall time of the step input.

%\subsection{Choosing a Varactor Diode}

\subsection{Methods of Modeling Varactors}

@@

\subsection{Modeling Board Characteristics}

@@

\subsubsection{SPICE Models}

@@

\subsubsection{Verilog-A Modules}

@@

\section{ Simulations }

@@

\subsection{Simulating Rise Time}

In our simulations we use an input step of 5 volts which takes $1\mu s$ to reach
it's final value. Using equation \ref{DelayApproxCalc}, we calculate a pulse
sharpening of $8.44ns$ at our $10^{th}$ node and $84.37ns$ at our $100^{th}$
node. To make these calculations, we used the CV plots obtained by measuring the
BBY40 with the test fixture discussed in section \ref{subsec:CVCurves}.

@@

For a lossless line, the sharpest pulse we can achieve can be estimated by first
using equation \ref{NLTLCutoff} taken from \cite{wilson1991pulse}. This gives us
the ideal cutoff frequency of the line which in Hz is about $f_c \approx
290Mhz$. The period of a wave moving at this rate is $T \approx 3.4ns$. While we
have not been able to achieve a rise time of this magnitude, we need to remind
ourselves of the inherent parasitics we are dealing with. Using a basic model
with values from our inductor and varactor diode datasheet. We have a per
section series resistance of roughly $R_s = 122m\Omega$ at DC, and a parallel
resistance given by the resistance inherent in the varactor packaging as well as
of the trace itself. This value was taken from the SPICE model used and was
$R_{p} = 1m\Omega$

>>>>

\section{ Theory }
\subsection{Why use Nonlinear Devices?}\label{sub:WhyNonLin}

@@

\subsection{Types of NLTL waves}\label{sub:NLTLTypes}
While treating the physics of non-linear waves is beyond the scope of this
paper, the use of NLTL's in transmission lines is founded upon the existence of
waves that do not behave like the pure sine waves we usually consider in
electronics. These both types of waves can be present and used in techonology to
lessen the time it takes for an input waveform to reach its final value as well
as lessen the time to return to its ``off'' voltage state. In other words, these
can be used to sharpen both rising and falling edges of a square
wave input\cite{afshari2005nonlinear}.

Informally, \emph{Solitons} are much like the type of wave we see
travelling in the open ocean without breaking and are characterized by their
ability to interact with other waves without losing energy. This means that
after passing another wave it largely maintains its shape and is only shifted in
phase (experiences time delay).

\emph{Shockwaves} are more easily visualized by a wave approaching the shore and
breaking. This is the type of wave we utilize when considering non-uniform
NLTL's. As the wave moves closer to the beach and the shoreline rises
dramatically, the wave can no longer move in the same way due to the shore
interrupting the circulating water and the momentum carried by it. With this
sudden discharge of energy, we see the wave ``break'' and momentarily create a
large vertical column of water.

\subsection{Non-linearity vs Wave Dispersion}\label{NonlinearVsDispersion}

Another main concept of operation of these types of circuits, is the need for
non-linearity to outweigh possible wave dispersion. Dispersion is a concept
commonly discussed relative to optical fiber cable as a limiting factor of
transmission speed. The idea is that travelling through any medium will cause a
small amount of pulse spreading. This means that, if dispersion is strong
enough, the pulse can spread or even break apart into several smaller pulses. To
overcome this, we need enough nonlinearity in the line primarily in terms of
voltage/amplitude dependence such that the wave remains narrow.

Figure \ref{fig:shockline}\cite{shocklineNLTL}, shows how construction of NLTL's
follows catch up theory and how pulse sharpening is achieved by these types of
circuits.

 \begin{figure}[htb]
\centering
\includegraphics[width=0.45\textwidth]{shockline_nltl_fig2.jpg}
\caption{\ This graphic taken from an ANRITSU application note shows the basic
principle of catch-up theory.
and how it can be used to sharpen input pulses}\label{fig:shockline}
\end{figure}

\subsection{Varactors vs. Nonlinear Capacitors}\label{sub:varacVSnonlincap}

@@

Equation (\ref{eq:LargeSignalDiodeCap}) becomes useful as we start looking at
larger pulses with large amplitude and should allow for scaling of our design.
\begin{equation}\label{eq:LargeSignalDiodeCap}
C_{ls}(V_{l},V_{n}) = \frac{1}{V_{n}-V_{l}}\int_{V_{l}}^{V_{n}}C(v)\, dv
\end{equation}

This leads to an important constraint when designing passive NLTL's made up of
series low-pass LC circuits. The fastest wave we can propagate down the NLTL can
be approximately found by using equation (\ref{eq:bragg}) to find the cutoff
frequency of the network as a function of the voltage across the varactors.

\begin{equation}\label{eq:bragg}
	f_{bragg} \approx \frac{1}{2\pi{}\sqrt{LC(v)}}	
\end{equation}

This equation is used in section \ref{sub:simsRise} indirectly to estimate the
amount of pulse sharpening we can expect for a given step inputl

\subsection{Methods of Modeling Varactors}\label{sub:VaracModeling}

@@

\subsection{Modeling Board Characteristics}\label{sub:BoardModeling}

@@

\subsubsection{SPICE Models}\label{subsub:SPICE}

@@

\subsubsection{Verilog-A Modules}\label{subsub:VerilogAMS}

@@

\section{ Simulations }\label{sec:sims}

@@

\subsection{Simulating Rise Time}\label{sub:simsRise}

In our simulations we use an input step of 5 volts which takes $1\mu s$ to reach
it's final value. Using equation (\ref{DelayApproxCalc}), we calculate a pulse
sharpening of $8.44ns$ at our $10^{th}$ node and $84.37ns$ at our $100^{th}$
node. To make these calculations, we used the CV plots obtained by measuring the
BBY40 with the test fixture discussed in section \ref{subsec:CVCurves}.

@@

For a lossless line, the sharpest pulse we can achieve can be estimated by first
using equation \ref{NLTLCutoff}\cite{wilson1991pulse}.
We notice that this utilizes the equation of our estimated bragg frequency as
seen in (\ref{eq:bragg}) and describes the differenece in cutoff
frequencies at maximim and minimum bias voltage multiplied by the number of
sections, N.

 This gives us the ideal cutoff frequency of the line which in Hz is about $f_c
 \approx 290Mhz$. The period of a wave moving at this rate is $T \approx 3.4ns$.
 While we have not been able to achieve a rise time of this magnitude, we need
 to remind ourselves of the inherent parasitics we are dealing with. Using a
 basic model
with values from our inductor and varactor diode datasheet. We have a per
section series resistance of roughly $R_s = 122m\Omega$ at DC, and a parallel
resistance given by the resistance inherent in the varactor packaging as well as
of the trace itself. This value was taken from the SPICE model used and was
$R_{p} = 1m\Omega$
